---
title: "SMML Class 1 Lab"
author: "Sunghee Lee"
date: "8/30/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{bbm}
- \usepackage{setspace}
- \singlespacing
number_sections: yes
fontsize: 12pt
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE)
```



## I. Made-up income data 1
```{r}
income1<-data.frame("id"=c(1:20),
                          "income_usd"=rnorm(20, mean = 10000, 
                                             sd = 10000/3), 
                          "educ_yr"=rep(12,20))
head(income1)
summary(income1)
```
1. How do income and education look?
2. Can you study the relationship between income and education?
12 is repeating, everything is same; therefore, we cannot study the relationship between income and education


## II. Made-up income data 2
```{r}
income2<-data.frame("id"=c(101:120),
                          "income_usd"=rep(10000, 20), 
                          "educ_yr"=sample(0:16,20,replace=T) )
head(income2)
summary(income2)
```
1. How do income and education look?
2. Can you study the relationship between income and education?
There is no relationship

## III. Wage data in R package ISLR
```{r, echo=FALSE}
library(ISLR)
```


```{r}
data("Wage")
#See p.13 of R manual_ISLR.pdf for data description
dim(Wage)
names(Wage)
head(Wage)
summary(Wage)
```
1. What do you observe?
  + What is the variable type for each variable? 
2. What stories would you like to study using this data?
3. How would you express those stories with formulas? 


## IV. psid data in R package faraway
```{r, echo=FALSE} 
library(faraway)
```

```{r}
data("psid")
#See p.78 of R manual_faraway.pdf for data description
head(psid)
summary(psid)
```
1. What do you observe?
+ What is the variable type for each variable? 
2. What stories would you like to study using this data?
3. How would you express those stories with formulas? 


## V. Fictitious data 
```{r, echo=FALSE} 
library(readxl)
```

```{r}
fic_dat7<-read_excel(
  "~/Desktop/Fictitious Data.xlsx",
  sheet=7)
head(fic_dat7)
#View(fic_dat7)
```

```{r}
#?read_excel
```

1. Compute the average of Distress_2019 and Distress_2021
```{r}
mean(fic_dat7$Distress_2019)
mean(fic_dat7$Distress_2021)
```
- Can you say that there is a difference in Distress between 2019 and 2022?

2. Compute the average of Distress_2019 and Distress_2021 by Sex
```{r}
cbind(aggregate(Distress_2019~Sex,fic_dat7,FUN=mean),
      aggregate(Distress_2021~Sex,fic_dat7,FUN=mean))
```
- Can you say that there is a difference in Distress between Male and Female in 2019? How about 2022?

## VI. Take home exercise

Data from the Health and Life Study of Koreans (HLSK) is available on Canvas, "HLSK.RDS". The codebook and other associated materials are available from https://www.icpsr.umich.edu/web/RCMD/studies/37635

1. Download the data from Canvas. You can read the data into R using the "readRDS" function.
2. Find household annual income variable. What difference do you see in this, compared to income in Wage and psid? 
3. What is the minumum, mean, mode, median and maximum of the income?
4. What is the variance and standard deviation of the income?
5. Visualize the income using a histogram and a box plot. Do you see the patterns in #3 and #4 in these? What are the benefits of each visualization method? How about drawbacks?
6. Going over the codebook and think about what kind of stories you want to learn about the income. How would you express those stories with formulas? 
 
