---
title: "SMML Class 1 Lab Take-Home"
author: "Sunghee Lee"
date: "8/30/2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{bbm}
- \usepackage{setspace}
- \singlespacing
number_sections: yes
fontsize: 12pt
fig_width: 7
fig_height: 6
---

 
```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE)
```


## Review. Take home exercise

Data from the Health and Life Study of Koreans (HLSK) is available on Canvas, "HLSK.RDS". The codebook and other associated materials are available from https://www.icpsr.umich.edu/web/RCMD/studies/37635

1. Download the data from Canvas. You can read the data into R using the "readRDS" function.
```{r}
HLSK<-readRDS(
  "C:/Users/sungh/Dropbox (University of Michigan)/2022 SMML 1/HLSK/HLSK.RDS")
#You need to change the path in the line above with yours as this 
#is unique to my computer
```

2. Find household annual income variable. What difference do you see in this, compared to income in Wage and psid? 
```{r}
#Check the codebook from ICPSR website: 37635-0001-Codebook-ICPSR.pdf
names(HLSK)
summary(HLSK$LQ3_PUB)
mean(HLSK$LQ3_PUB, na.rm=T)
mn<-function(x){
  mean(x, na.rm=T)
  }
mn(HLSK$LQ3_PUB)
```

3. What is the minimum, mean, mode, median and maximum of the income?
```{r}
summary(HLSK$LQ3_PUB)

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
  }

Mode(HLSK$LQ3_PUB)
```

4. What is the variance and standard deviation of the income?
```{r}
var(HLSK$LQ3_PUB, na.rm=T)
sd(HLSK$LQ3_PUB, na.rm=T)
sqrt(var(HLSK$LQ3_PUB, na.rm=T))
```

5. Visualize the income using a histogram and a box plot. Do you see the patterns in #3 and #4 in these? What are the benefits of each visualization method? How about drawbacks?
```{r} 
library(ggplot2)
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_histogram()
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_histogram(binwidth=1000)
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_histogram(binwidth=5000)
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_histogram(binwidth=10000)

ggplot(HLSK, aes(x=LQ3_PUB)) + geom_boxplot()
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_boxplot(notch=TRUE)
ggplot(HLSK, aes(x=LQ3_PUB)) + geom_boxplot(notch=TRUE) + 
  geom_dotplot(binaxis='x', stackdir='center', dotsize=0.2)

```

6. Going over the codebook and think about what kind of stories you want to learn about the income. How would you express those stories with formulas? 
* What are potential factors that may influence immigrants' income ($y_i$) given the HLSK data?
  * Examples
    * Time of immirgration (e.g., before or after age 18)
    * Country where the final degree was obtained (e.g., US. vs. non-US)
    * Years in the U.S.
* Is having a final degree from the US associated with higher income for Korean immigrants than a degree from elsewhere?
  * Formula: $\mu_{US}>\mu_{Non-US}$
* What is the relationship between years in the U.S. and immigrants' income?
  * Formula: $\rho_{income,USyrs}$
* What is the effect of one more year in the U.S. on immigrants' income?
  * Formula: $y_{i}=\beta_0+\beta_1x_i+\epsilon_i$
 

